Spatial evolution of solar wind discontinuities in the outer heliosphere

### 2024-09-23 article.qmd
update: occurrence rate analysis and consider different hypotheses

@sodingRadialLatitudinalDependencies2001
@leppingMagneticFieldDirectional1986
@marianiStatisticalStudyMagnetohydrodynamic1983
@marianiVariationsOccurrenceRate1973
@tsurutaniInterplanetaryDiscontinuitiesTemporal1979

TODO: how is the occurrence rate related to the formation mechanism?

>>>>

The observed increase in the average waiting time between discontinuities with radial distance prompts us to explore different hypotheses to explain this phenomenon. At any given radial distance, discontinuities may originate from two primary sources: those generated near the Sun and convected outward by the solar wind, and those formed locally. Since the solar wind speed is significantly higher than the Alfvén speed—which characterizes the propagation speed of rotational discontinuities (RDs)—the fraction of inward-propagating discontinuities is expected to be negligible. Additionally, when observations are made by a single spacecraft, some discontinuities may be missed due to unfavorable positioning or limitations in detection methods.

First, we consider the limitations of our current detection methods. An increase in the duration of discontinuities could render some of them unidentifiable if their duration becomes comparable to the time interval used in our analysis. Although we observe a slight increase in the average duration of discontinuities with radial distance, this increase is relatively small. The overall population of discontinuities is still predominantly characterized by those with short durations (see Figure X in the appendix). Since our method normalizes the change in the magnetic field by the surrounding magnetic field, a decrease in magnetic field variation with radial distance alone does not necessitate non-identification. Notably, our results differ from previous studies [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979], which reported typical durations of 2 to 3 seconds at 1 AU and 10 to 30 seconds at 5 AU. In our analysis, however, the observed durations remain below 10 seconds at 5 AU, suggesting that our method may be more effective at identifying shorter-duration events than those reported in prior studies.

If the majority of distant discontinuities are generated locally, and assuming that the primary conditions for their generation do not change significantly with radial distance, we would expect them to be uniformly distributed throughout the hemisphere. Consequently, the occurrence rate should be independent of radial distance. To test this hypothesis, we examined the relationship between waiting time and plasma beta. However, even at a 99% statistical significance level, we could not reject the null hypothesis that these two variables are independent. Therefore, if discontinuities are primarily generated locally, there must be other factors influencing their generation rates.

Alternatively, we consider the hypothesis that the majority of discontinuities are generated close to the Sun. If we assume that discontinuities are one-dimensional, infinitely extended structures, then without annihilation, all discontinuities passing through one location would also pass through other radial distances. However, a more realistic assumption is that discontinuities are three-dimensional curved surfaces with lengths and widths proportional to their thickness. In this scenario, the occurrence rate observed by a single spacecraft does not represent the total number of discontinuities within a given shell. Since our spacecraft travels within the ecliptic plane and the normal direction of most discontinuities lies within this plane, it is necessary to project the discontinuities onto a circular cross-section. The projected lengths should then be normalized by the circumference of the circle to estimate the total number of discontinuities within the cylindrical shell. The normalization factor is  $d \sin\theta_{n,v} / r$ , where $d$ is the thickness of the discontinuity, $\theta_{n,v}$ is the angle between the normal direction of the discontinuity and the solar wind velocity, and  $r$ is the radial distance from the Sun (the result is shown in the appendix). Accounting for this factor reduces the differences in the normalized occurrence rates observed by different spacecraft. However, we cannot rule out the possibility that some discontinuities may be annihilated as they convect outward with the solar wind. This annihilation process would result in a lower number of discontinuities observed by Juno at larger radial distances. Therefore, under this hypothesis, the annihilation process must play a significant role. (However, if we disregard changes in the thickness and only account for the factor of the circumference of the circle, the normalized discontinuities observed by Juno would be consistent with those observed at 1 AU.)

![](figures/ocr/ocr_normalized_by_r.svg)
